### 2024-01-16 302tut2 code.qmd
Add files via upload

---
title: "minipaper2-opendatatoronto"
author: "Yichen Ji (Selina)"
format: html
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
#install.packages("opendatatoronto")
```

```{r}
library(dplyr)
library(knitr)
library(opendatatoronto)
library(tidyverse)
library(purrr)
library(ggplot2)
packages <- list_packages(limit = 10)

ttc_packages <- search_packages("ttc")

ttc_delay_packages <- search_packages("TTC Subway Delay")

ttc_delay_resources <- ttc_delay_packages %>%
  list_package_resources()

ttc_delay_statistics <- ttc_delay_resources %>%
  slice(5) %>%
  get_resource()

ttc_tibble <- map_dfr(ttc_delay_statistics, ~tibble(
  station = .x$Station,
  code = .x$Code,
  min_delay = .x$`Min Delay`
))

ttc_tibble

ttc_tibble%>%filter(min_delay > 0)

station_delays <- ttc_tibble%>%
  group_by(code) %>%
  summarise(Delay_Count = n())%>%
  ungroup() %>%
  top_n(10, Delay_Count) %>%
  arrange(Delay_Count)

# Plot the histogram
ggplot(station_delays, aes(x = reorder(code, Delay_Count), y = Delay_Count)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Number of Delays per Station", x = "Station", y = "Number of Delays")
```