### 2024-01-18 outputs/paper/paper.qmd
Change

---
title: "What influences the Death "
subtitle: "My subtitle if needed"
author: 
  - Yichen Ji
thanks: "Code and data are available at: https://github.com/Selinayichenji/2024-Data-collection.git."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r, echo = FALSE}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false
#install.packages("scales")
library(knitr)
library(scales)
library(kableExtra)

total_individuals <- nrow(census_data)

children_number_census <- census_data %>%
  group_by(gender) %>%
  summarise(
    Total_Number = n(),
    Ratio_to_Total = percent(n()/total_individuals, accuracy = 0.01)
  ) %>%
  ungroup()

children_number_census %>%
  kable("latex", booktabs = TRUE, caption = "Summary of Children Number Groups") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References