### 2024-01-25 outputs/paper/paper.qmd
final paper

---
title: "Investigating the Effect of COVID-19 to the Death of Homeless People in Toronto from 2017 to 2023"
subtitle: "Examining Trends in Age, Gender, and Causes of Death Among Toronto's Homeless Population from 2017 to 2023"
author: 
  - Yichen Ji
thanks: "Code and data are available at: https://github.com/Selinayichenji/2024-Data-collection.git."
date: "`r format(Sys.time(), '%d %B %Y')`"
date-format: long
abstract: "This study investigated mortality trends among Toronto's homeless population from 2017 to 2023. The findings suggest that the primary contributor to the elevated mortality rate during the pandemic was drug toxicity, rather than COVID-19. The objective of this research is to accurately identify the pressing issues and facilitate effective assistance for the homeless community."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
#code referenced from posit community: https://community.rstudio.com/t/force-table-placement/149954
header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}

---

```{r setup, include = FALSE}

#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(ggplot2)
#print(citation("tidyr"), bibtex=TRUE)
analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
```

# Introduction

In recent decades, Toronto has witnessed an increasingly visible homelessness issue, with individuals seeking refuge on city streets becoming a common sight. In 2013, homeless people took 0.19% [@cite13data] of total population in Toronto, ranked the 8th city in Canada. In 2021, there was 7,347 [@citehh] homeless people in toronto and the total population was 2,794,356 [@citehh]. The percentage rose to 0.26%. This trend not only highlights a growing societal challenge but also underscores an urgent need for action.

As [@cite13data] said, “The goal of ending homelessness is to ensure housing stability...and includes required services as needed (supportive), in addition to income and supports.” In this paper, we are trying to figure out what caused the increase death of homeless people during pandemic in order to provide social service they needed by using R [@citeR] to analysis a dataset from opendatatoronto [@citeopendatatoronto], named "Homeless deaths by cause".

And there are some limitations of this data listed on the City of Toronto Website we must know before using the data:
"1. This data reflects only deaths that are reported to TPH by SSHA, community partners and the Coroner’s Office.
2. Cause of death is unknown or pending in approximately 25% of the reported deaths.
3. In 2019, TPH ceased reporting on location of death as this information is often unknown or unverified.
4. The data does not identify Indigenous status as this is reported as unknown or missing in 70% of the reported cases. With this high a percentage of missing information, accurate conclusions cannot be drawn, as such, it is standard epidemiological practice to suppress the release of data.
5. Cause of death for transgender people not shown due to small counts.
6. Cause of death is unknown in approximately 25% of the reports. To protect privacy, causes of death with less than 2% of the cases are included in the other category." [@citecot]

Through tables, plot graphs and bar graphs, our findings reveal a huge increase in deaths attributed to drug use during the pandemic, overshadowing even COVID-19 as a direct cause of death. This suggests a secondary crisis unfolding within the pandemic period.The precise reasons behind this appearance remain ambiguous and subject to further professional research, several hypotheses have been proposed. Dr.Hwang observed, '“The disruption caused by the lockdown and by the real lack of availability of services caused people to change where they spend time and thus where they use drugs,” ... “I think that contributed to people using in higher risk situations that would more likely result in death.”'[@citenews]. The experience may prove beneficial for the charity's work with homeless individuals, particularly when critical times such as a pandemic come in the future.

# Data {#sec-data}

## Data source and background

The data used through this paper was provided by the City of Toronto's OpenDataToronto Library, was a cvs file named "Homeless deaths by cause", under the category "Deaths of People Experiencing Homelessness". The homelessness is defined as “the situation of an individual or family without stable, permanent, appropriate housing, or the immediate prospect, means and ability of acquiring it”[@citecot]. We retrieved the data from R package opendatatoronto[@citeopendatatoronto].We downloaded the data, cleaned it and renamed it as "analysis_data" in output/datam folder.

The data was measured and collected by Toronto Public Health (TPH), Support and Housing Administration (SSHA) and the Office of the Chief Coroner of Ontario (OCCO) since January 2017. The TPH tracked the deaths of homeless people to get accurate estimate of their number and causes of death. The TPH also leads the data collection, analysis and reporting. The SSHA and health and social service agencies that support homeless people share information about a death with TPH and the OCCO verifies some of the data. [@citecot]

## Variables of interest

Our population is divided by 4 categories: Year of death, Gender, Age group and Cause of death. The year of death covers form 2017 to 2023. The gender category has 3 types: Female, Male and Unknown. Age groups are divided as below: Under 20, 20-39,40-59,60+ and Unknown.Causes of death includes 10 types below: Accident, Cancer, Cardiovascular Disease, COVID-19, Drug Toxicity, Homicide, Other, Pneumonia, Suicide and Unknown/Pending.

## Data processing

We used `R`[@citeR] to process the whole paper, packages include `tidyverse`[@citetdv], `dplyr`[@citedpr], `tidyr`[@citetdr], `knitr`[@citeknr], `kableExtra`[@citekable] and `ggplot2`[@citegg2].

## Summarized analysis data

Then there are basic distributions of population in gender, age and causes of death.

```{r}
#| label: tbl-gender-organize
#| tbl-cap: Gender Distribution by Year
#| echo: false
#| warning: false
#| message: false

# Calculate the yearly gender counts
yearly_gender_counts <- analysis_data %>%
  group_by(year_of_death, gender) %>%
  summarise(count = n(), .groups = 'drop')

# Reshape to a wide format
gender_data_wide <- yearly_gender_counts %>%
  pivot_wider(names_from = gender, values_from = count, values_fill = list(count = 0))

# Add a sum row for each year
gender_data_wide <- gender_data_wide %>%
  rowwise() %>%
  mutate(Sum = sum(c(Male, Female, Unknown), na.rm = TRUE))

# Calculate the sum of every row
totals <- colSums(gender_data_wide[, c("Female", "Male", "Unknown", "Sum")], na.rm = TRUE)

totals <- as.list(totals)
gender_data_wide$year_of_death <- as.character(gender_data_wide$year_of_death)

final_data <- gender_data_wide %>%
  add_row(year_of_death = "Total", Female = totals$Female, Male = totals$Male, Unknown = totals$Unknown, Sum = totals$Sum)

# Render final data with kable
final_data %>%
  kable(format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped"))

```

The referenced table @tbl-gender-organize shows gender-specific mortality among the homeless over seven years. The 'Sum' column totals annual deaths, with males constituting 731 cases, or 78% of all deaths. Females accounted for about a third of male deaths, and the 'Unknown' category was minor, with 12 cases.

```{r}
#| label: tbl-age-organize
#| tbl-cap: Age Distribution by Year
#| echo: false
#| warning: false
#| message: false
# Calculate the yearly gender counts
yearly_age_counts <- analysis_data %>%
  group_by(year_of_death, age_group) %>%
  summarise(count = n(), .groups = 'drop')

# Reshape to a wide format
age_data_wide <- yearly_age_counts %>%
  pivot_wider(names_from = age_group, values_from = count, values_fill = list(count = 0))

# Calculate total amount of each gender
total_age_counts <- analysis_data %>%
  group_by(age_group) %>%
  summarise(Total = n(), .groups = 'drop')

# Switch row and column
total_age_wide <- total_age_counts %>%
  pivot_wider(names_from = age_group, values_from = Total, values_fill = list(Total = 0))

# Add a 'Total' identifier in the year_of_death column for joining
total_age_wide <- total_age_wide %>%
  mutate(year_of_death = "Total")

#Ensure we combine the same type row
age_data_wide$year_of_death <- as.character(age_data_wide$year_of_death)
total_age_wide$year_of_death <- as.character(total_age_wide$year_of_death)

# Combine the yearly data with the totals
final_data <- bind_rows(age_data_wide, total_age_wide)

final_data %>%
  kable(format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped"))

```

@tbl-age-organize reveals the 40-59 age bracket had the highest mortality at 387, while under-20s had the fewest at 9. The "20-39" and "60+" groups had 240 and 260 deaths, respectively, peaking in 2021 for the former and 2022 for the latter. The under-20 category had 3 deaths in both 2021 and 2022.

A seven-year trend from @tbl-gender-organize shows a dip in deaths from 2017 to 2018, a surge through 2022, and a decline to below 80 in 2023. This period coincides with the COVID-19 pandemic, suggesting a potential link to homeless mortality; however, age and gender had minimal impact on death numbers.

```{r}
#| label: tbl-cause-organize
#| tbl-cap: Death Causes Distribution by Year - first half
#| echo: false
#| warning: false
#| message: false

death_cause_counts <- analysis_data %>%
  group_by(year_of_death, cause_of_death) %>%
  summarise(count = n(), .groups = 'drop')

cause_data_wide <- death_cause_counts %>%
  pivot_wider(names_from = cause_of_death, values_from = count, values_fill = list(count = 0))

total_cause_counts <- analysis_data %>%
  group_by(cause_of_death) %>%
  summarise(Total = n(), .groups = 'drop')
total_cause_wide <- total_cause_counts %>%
  pivot_wider(names_from = cause_of_death, values_from = Total, values_fill = list(Total = 0))

total_cause_wide <- total_cause_wide %>%
  mutate(year_of_death = "Total")

#Ensure we combine the same type row
cause_data_wide$year_of_death <- as.character(cause_data_wide$year_of_death)
total_cause_wide$year_of_death <- as.character(total_cause_wide$year_of_death)

# Combine the yearly data with the totals
final_data_2 <- bind_rows(cause_data_wide, total_cause_wide)

# Divide our format into 2 parts in order to ensure the content fits within the PDF page.
# Divide at the fifth column
first_half <- final_data_2[, 1:5]

# Build format for each part
first_half %>%
  kable(format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-cause-organize-2
#| tbl-cap: Death Causes Distribution by Year - second half
#| echo: false
#| warning: false
#| message: false
second_half <- final_data_2[, c(1, 6:ncol(final_data_2))]
second_half %>%
  kable(format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped"))

```

Drug toxicity is the leading death cause, as shown in @tbl-cause-organize and @tbl-cause-organize-2, accounting for half of the deaths, peaking in 2021. Unknown causes and cardiovascular diseases followed, the latter spiking to 22 in 2022 and then falling to 1 in 2023. COVID-19 related deaths were the lowest at 8, with half in 2020. Other causes ranged between 20 to 60 deaths.

The data suggest COVID-19 indirectly influenced mortality through increased drug toxicity and cardiovascular deaths rather than direct impact. Further analysis of the interplay between gender, age, and mortality causes is needed to substantiate secondary hypotheses.

# Data Analysis {#sec-ana}

## Analysis on Causes of Death Based on Gender

```{r}
#| label: fig-gender-in-deathcauses
#| fig-cap: Correlation between gender and 10 causes of death
#| echo: false
#| warning: false
#| message: false
gender_death <- analysis_data %>%
  ggplot(mapping = aes(x = cause_of_death, fill = gender)) +
  geom_bar() +
  labs(x = "Causes of death", y = "Number of observations", fill = "Gender") +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 10),
        axis.text.x = element_text(size = 7, angle = 45, hjust = 1))

gender_death
```

@fig-gender-in-deathcauses illustrates that across ten death causes, gender distribution is fairly consistent, with drug toxicity as the primary cause and COVID-19 the least for all genders. Female deaths are about a third of male deaths, aligning with data from @tbl-gender-organize.

## Analysis on Causes of Death Based on Age

```{r}
#| label: fig-age-in-deathcauses
#| fig-cap: Correlation between age and 10 causes of death
#| echo: false
#| warning: false
#| message: false
age_death <- analysis_data %>%
  ggplot(mapping = aes(x = age_group, fill = cause_of_death)) +
  geom_bar() +
  scale_fill_discrete(labels = c("Accident", "Cancer", "Cardiovascular\nDisease", "COVID-19","Drug Toxicity", "Homicide", "Other", "Pneumonia", "Suicide", "Unknown/\nPending")) +
  labs(x = "Age group", y = "Number of observations", fill = "Age group") +
  theme(legend.position = "bottom", #set labels under the figure
        legend.key.size = unit(0.5, 'cm'),
        legend.title = element_text(size = 10),# Adjust labels' size
        legend.text = element_text(size = 7)) 
age_death

```

Age-related death causes, shown in @fig-age-in-deathcauses, indicate that the under-20 group is too small for analysis, and the unknown group lacks clear trends. The focus thus is on the 20-39, 40-59, and 60+ groups. Drug toxicity dominates in the 20-59 brackets, while unspecified causes, cancer, and cardiovascular diseases are more common in the 60+ group, with drug-related deaths less prevalent.

\newpage

## Analysis of the Distribution of Causes of Death Across Three Age Groups

```{r}
#| label: fig-20-39group-deathcause-change
#| fig-cap: Trends in causes of death among age group "20-39"
#| echo: false
#| warning: false
#| message: false

above20_causes <- analysis_data %>%
  filter(age_group == "20-39") %>%
  group_by(year_of_death, cause_of_death) %>%
  summarise(deaths = n(), .groups = 'drop')

above20_time_plot <- ggplot(above20_causes, aes(x = year_of_death, y = deaths, color = cause_of_death, group = cause_of_death)) +
  geom_line() +
  scale_y_continuous(limits = c(0, NA))+
  scale_color_discrete(labels = c("Accident", "Cancer", "Cardiovascular\nDisease",
                                  "Drug Toxicity", "Homicide", "Other", "Pneumonia",
                                  "Suicide", "Unknown/\nPending")) +
  labs(x = "Year of Death",
       y = "Number of Deaths",
       color = "Cause of Death") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 10),)

above20_time_plot 
```

```{r}
#| label: fig-40-59group-deathcause-change
#| fig-cap: Trends in causes of death among age group "40-59"
#| echo: false
#| warning: false
#| message: false

above40_causes <- analysis_data %>%
  filter(age_group == "40-59") %>%
  group_by(year_of_death, cause_of_death) %>%
  summarise(deaths = n(), .groups = 'drop')

above40_time_plot <- ggplot(above40_causes, aes(x = year_of_death, y = deaths, color = cause_of_death, group = cause_of_death)) +
  geom_line() +
  scale_y_continuous(limits = c(0, NA))+
  scale_color_discrete(labels = c("Accident", "Cancer", "Cardiovascular\nDisease", "COVID-19",
                                  "Drug Toxicity", "Homicide", "Other", "Pneumonia",
                                  "Suicide", "Unknown/\nPending")) +
  labs(x = "Year of Death",
       y = "Number of Deaths",
       color = "Cause of Death") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 9))

above40_time_plot
```

```{r}
#| label: fig-60group-deathcause-change
#| fig-cap: Trends in causes of death among age group "60+"
#| echo: false
#| warning: false
#| message: false

plus60_causes <- analysis_data %>%
  filter(age_group == "60+") %>%
  group_by(year_of_death, cause_of_death) %>%
  summarise(deaths = n(), .groups = 'drop')

plus60_time_plot <- ggplot(plus60_causes , aes(x = year_of_death, y = deaths, color = cause_of_death, group = cause_of_death)) +
  geom_line() +
  scale_y_continuous(limits = c(0, NA))+
  scale_color_discrete(labels = c("Accident", "Cancer", "Cardiovascular\nDisease", "COVID-19",
                                  "Drug Toxicity", "Homicide", "Other", "Pneumonia",
                                  "Suicide", "Unknown/\nPending")) +
  labs(x = "Year of Death",
       y = "Number of Deaths",
       color = "Cause of Death") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 9))

plus60_time_plot
```

Age shows a stronger link to death causes than gender, prompting a shift to analyze temporal trends in these age groups. Trends indicate a surge in drug toxicity deaths from 2017 to 2021, especially in the 20-39 and 40-59 groups, with a decline afterward. The 40-59 group also shows a notable number of deaths from unknown causes.

For the over-60s, as per @fig-60group-deathcause-change, drug toxicity and unspecified deaths dipped in 2021 but rebounded in 2022, with cancer deaths peaking that year, forming an "M"-shaped trend. Cardiovascular deaths have increased, peaking in 2022, while other causes remain below five.

\newpage

# Conclusion {#sec-conclu}

In conclusion, our data analysis uncovers a distinct trend and points to potential factors behind the increase in drug toxicity-related deaths among Toronto's homeless population from 2017 to 2023. The evidence, drawn from @tbl-gender-organize, @fig-20-39group-deathcause-change, @fig-40-59group-deathcause-change, and @fig-60group-deathcause-change, shows a sharp rise in death among the young and middle-aged homeless segments. In contrast, the older homeless population appears to be less affected by drug-related issues, with cardiovascular disease being their leading cause of mortality. While our data does not provide evidence of the underlying causes of these trends, experts have speculated that the lockdowns may have compelled individuals to use drugs in risker and stranger circumstances, potentially leading to higher death number related to drug toxicity. [@citenews]

# Appendix {#sec-app}

We cleaned the data by cancelling the Count	column, which represents the number of deaths in the specified category[@citecot]. For the convenience of counting specific population group, we copied the particular line a couple of times according to the number in the count. And all characters in columns' names were transformed into lowcase.

Here is the sample of cleaned data.

```{r}
#| label: tbl-glimpse
#| tbl-cap: Sample of Data
#| echo: false
#| warning: false
#| message: false

sample_data<-head(analysis_data, 10)
kable(sample_data, format = "latex", booktabs = TRUE)%>%
  kable_styling(position = 'center')
```

\newpage

# References